### 2020-04-21 final_report/data.tex
add reference to dimensional reduction
add section on autoencoders

Since most surrogate models
that we employ assume overall continuous inputs, we take steps to unify our
feature interface in order to attain this property. In particular, we eliminate
discrete features by embedding each such feature into a finite multitude of
continuous features using standard one-hot encoding. This option is available to
us since discrete domains that generate our feature space are finite in
cardinality and relatively small in size. And while it renders all features continuous, the
modification comes at the expense of increasing the dimensionality of the
feature space to 27 features in total.

@@

** TODO: Make sure discrete slices are mentioned before this **

@@

TODO: Question: is it accurate to say this represents nonlinear feature extraction? re my intro paragraph

TODO
\begin{enumerate}
	\item working principle
	\item application on single slice
	\item application on mixed slice data
\end{enumerate}

>>>>

Since most surrogate models
that we employ assume overall continuous inputs, we take steps to unify our
feature interface in order to attain this property. In particular, we eliminate
discrete features by embedding each such feature into a finite multitude of
continuous features using standard one-hot encoding. This option is available to
us since discrete domains that generate our feature space are finite in
cardinality and relatively small in size. And while it renders all features continuous, the
modification comes at the expense of increasing the dimensionality of the
feature space to 27 features in total. This is further discussed
in~\cref{sec:dimred}.

@@

@@

\begin{wrapfigure}{r}{0.25\textwidth}
	\centering
	\vspace{-3ex}
	{\footnotesize \incfigscale{0.8}{autoencoder}}
	\caption{Autoencoder with input width~$W_0$ and bottleneck width~$W_1$.}
	\label{fig:autoencoder}
\end{wrapfigure}

Autoencoders are a family of approaches to feature extraction driven by neural
networks. Faced with many possible alternatives, we implemented a conventional
autoencoder, which relies on a three-layer network that is trained to replicate
the identity mapping. While it follows that the input and output layers of such
network are sized to accommodate the analysed dataset, the
hidden layer---also called the bottleneck---allows for variable number of
neurons that represent a smaller subspace. By scanning
over a range of bottleneck widths and investigating relative changes in the
validation loss, we assess the potential for dimensional reduction.

This approach was applied to two sets of samples: (i) a subset of data obtained
from run~1 and (ii) a subset of a single slice obtained from run~2. Our
expectation was that while the former case would provide meaningful insight into
correlations within the feature space, the latter would validate our
autoencoder implementation by analysing a set of points that are trivially
reducible due to a number of fixed features.

\begin{wrapfigure}{l}{0.4\textwidth}
	\centering
	\vspace{-3ex}
	% TODO: placeholders only, regenerate when time is available
	\begin{subfigure}[b]{\linewidth}
		\centering
		\includegraphics[width=\linewidth]{ae_plot_all_slices_combined}
	\end{subfigure}
	\begin{subfigure}[b]{\linewidth}
		\centering
		\includegraphics[width=\linewidth]{ae_plot_single_slice}
	\end{subfigure}
	\caption{Autoencoder loss in two considered cases of dimensional reduction.}
	\label{fig:autoencoder-loss}
\end{wrapfigure}

The results of both experiments are shown in~\cref{fig:autoencoder-loss}.
Consistent with our motivation, in each plot we can clearly identify a constant
plateau in the region of large dimensionality followed by a point, from which a
steep increase in loss is observed towards lower dimension counts. We consider
this critical point to correspond with the largest viable dimensional reduction
without significant information loss.
With this methodology we find that the autoencoder was able to reduce the
datasets into a subspace of 17~dimensions in the first case and 10~dimensions in
the second case.

Confirming our expectation that in the latter, trivially reducible case the
autoencoder should achieve stricter dimensional reduction, we are
encouraged to believe that our implementation is indeed efficient. However, we
must also conclude that in both examined cases this method failed to produce a
reduction that would be dominate a naïve approach, in which we eliminate
7~dimensions due to overdetermined one-hot-encoded features, 2~dimensions due to
sum-to-one constraints, and in the latter case 7~additional dimensions due to
artificially fixed features. This supports the findings from PCA as well as the
presented semivariograms.

% TODO: Question: is it accurate to say this represents nonlinear feature extraction? re my intro paragraph
